Urban Crime Dynamics: Trends in Theft and Assault Over 2014 to 2023

### 2024-09-27 paper/paper.qmd
change the data analyse

Between 2014 and 2023, as urbanization accelerates, the crime dynamics in Toronto change significantly. Crime is concentrated in certain neighborhoods, such as those with poor socioeconomic conditions, where residents tend to face a higher risk of property and violence [@wang2019spatial]. Since 2020, the Covid-19 outbreak has intensified this trend. [@syamsuddin2021effect] noted a spike in thefts in many cities around the world during the pandemic, due to socio-economic instability and reduced social activity caused by lockdowns. Toronto's neighbourhoods are no exception, with economic pressures leading to a marked increase in property crime in some neighbourhoods. In addition, population growth also increases the chance of crime, and the chance of crime is higher in areas with higher population density and resource shortage [@bettencourt2007growth].

Data on Neighbourhood Crime Rates were obtained, as described in @sec-data-overview. This paper mainly analyzes the burglary and assault rates in each community from 2014 to 2023, exploring the differences in community size and crime rates and the accompanying time fluctuations. The research focuses on the dynamic change of crime rate in different communities before and after the epidemic (@fig-theft-rate-trend-2014-to-2023 and @fig-Box-plot-of-theft-rates) and the relationship between population density and crime rates in 2023 (@fig-population-and-assault-rate and @fig-Bar-chart-2023-theft-rates). By assessing these dynamics, we hope to inform crime prevention and resource allocation in Toronto, particularly in high-risk neighbourhoods where there is a long-term need for more police and social resources.

@@

@fig-theft-rate-trend-2014-to-2023 shows trends in the Top 15 and Bottom 15 Cities by Average Theft Rate from 2014 to 2023. Regardless of the theft rate, most areas show relatively stable crime rate in a long period of time, which may prove that this area has relatively perfect social security management mode and relatively stable life state. In addition, crime rates vary widely in some areas and are generally higher than in others. It is worth noting that between 2019 and 2020, most regions showed a trend of rising theft rates, reflecting changes in social conditions or economic development at the time. Overall, theft rates remained relatively flat in most regions, with only slight fluctuations. In addition, there are significant differences in the theft rate between different regions during the same period, which may be related to different social conditions in different regions, such as law enforcement efforts, economic conditions, population distribution and other factors.

@@

The overall distribution of theft rates and their outliers for each region between 2014 and 2023 can be studied using @fig-Box-plot-of-theft-rates. First, while the median showed a relatively flat change, the median theft rate increased slightly between 2018 and 2023. In 2020 and 2021 in particular, the median appears to have reached a higher position. In addition, outliers (indicated by black dots) show unusually high theft rates across years. It is worth mentioning that more significant outliers have appeared in 2019, and the number of outliers has increased significantly. This indicates that in 2019 there were a large number of areas with much higher than general theft rates, and that this was the case in multiple areas. Secondly, the quartile distance also showed a certain fluctuation in this year. The reasons for this may be related to changes in the socio-economic environment during the year, changes in enforcement policies, or social unrest.

@@

@fig-population-and-assault-rate shows that there is no obvious linear relationship between population and attack rates across regions in 2023. However, it can be found that the attack rate in the more populated areas (close to 30,000 people) seems to be relatively high and has an increasing trend, while the attack rate in the less populated areas is more volatile. And there are several outliers that show very high attack rates in areas with higher populations.

@@

This bar chart (@fig-Bar-chart-2023-theft-rates) shows the 20 areas with the highest theft rates in Toronto in 2023. The chart shows that the Yonge-Bay Corridor has a significantly higher theft rate than all other areas, at 312.8. Several other areas with high theft rates, such as Humber Summit and Yorkdale-Glen Park, also had relatively high theft rates of 255 and 235.5, respectively. This suggests that the area may have different conditions or characteristics than other areas, such as high traffic or rich commercial clusters that attract thieves.

>>>>

Between 2014 and 2023, as urbanization accelerates, the crime dynamics in Toronto change significantly. Crime is concentrated in specific communities, such as those with poor socio-economic conditions. In some marginalized communities in Toronto, due to deteriorating economic conditions and insufficient social resources, the crime rate is always high, and residents in this community are often at higher risk of property and violent crime, such as theft and assault [@wang2019spatial].

Since 2020, the Covid-19 outbreak has intensified this trend. Thefts have spiked in many cities around the world during the pandemic due to economic instability worldwide and reduced social interaction due to lockdown policies [@syamsuddin2021effect]. Toronto's neighbourhoods are no exception, with economic pressures leading to a marked increase in property crime in some neighbourhoods. In addition, population growth is also a factor in the rising crime rate. Population growth increases the opportunity for crime, and areas with higher population density and tight resources have more opportunities for crime [@bettencourt2007growth]. That's because densely populated areas have limited law enforcement resources, and they can't respond to crime suspects everywhere in a short period of time.

In order to further explore the spatio-temporal changes of the crime rate in Toronto, this study obtains the crime data of various communities from 2014 to 2023 according to the method described in @sec-data-overview, especially the data of theft and assault rates. Through these data, this paper analyzes the change trend of crime rates in different communities during these years. The differences between community size and crime rate and the fluctuations in different time periods are discussed. In addition, The research focuses on the dynamic change of crime rate in different communities before and after the epidemic (@fig-theft-rate-trend-2014-to-2023 and @fig-Box-plot-of-theft-rates) and the relationship between population density and crime rates in 2023 (@fig-population-and-assault-rate and @fig-Bar-chart-2023-theft-rates). By assessing these dynamics, this paper aims to inform law enforcement and city planners in Toronto, particularly with regard to resource allocation and crime prevention in high-risk neighbourhoods.

@@

The @fig-theft-rate-trend-2014-to-2023 shows trends for the top 15 and bottom 15 cities with average theft rates from 2014 to 2023. Regardless of the theft rate, the crime rate in most areas is relatively stable over a long period of time, which may prove that the social security management model in the area is relatively perfect, and the living condition is relatively stable. In addition, crime rates vary widely in some areas and are generally higher than in others. Notably, most regions showed a trend of rising theft rates between 2019 and 2020, which is related to the worsening socio-economic conditions caused by the Covid-19 outbreak at the time studied. [@syamsuddin2021effect]. Overall, theft rates remained relatively flat in most areas, with only slight fluctuations. In addition, there are clear differences in theft rates between different regions over the same period, which may be related to different social conditions in different regions, such as more opportunities for crime in areas with higher population density and tight resources [@bettencourt2007growth]. We will discuss the relationship between population and crime rates later on @fig-population-and-assault-rate and @fig-Bar-chart-2023-theft-rates.

@@

The overall distribution of theft rates and their outliers for each region between 2014 and 2023 can be studied using @fig-Box-plot-of-theft-rates. First, while the median showed a relatively flat change, the median theft rate increased slightly between 2018 and 2023. In 2020 and 2021 in particular, the median appears to have reached a higher position. In addition, outliers (indicated by black dots) show unusually high theft rates across years. It is worth mentioning that more significant outliers have appeared in 2019, and the number of outliers has increased significantly. This indicates that in 2019 there were a large number of areas with much higher than general theft rates, and that this was the case in multiple areas. Secondly, the quartile distance also showed a certain fluctuation in this year. This is evidence of the abnormal fluctuation in crime rates caused by Covid-19 [@syamsuddin2021effect].

@@

The @fig-population-and-assault-rate shows that there is no significant linear relationship between population and assault rates by region in 2023. However, it can be found that the assault rate in more populated areas (close to 30,000 people) appears to be relatively high and on the rise, while the fluctuation is greater in less populated areas. And there are several outliers that show very high rates of assault in more populated areas. This is consistent with the point made by Bettencourt et al. (2007) that areas with larger populations tend to experience higher crime rates due to increased social interaction and strained resources [@bettencourt2007growth].

@@

This bar chart (@fig-Bar-chart-2023-theft-rates) shows the 20 areas in Toronto with the highest theft rates in 2023. The chart shows that the Yonge-Bay Corridor area has a significantly higher theft rate than all other areas, at 312.8. Far beyond his several areas with high theft rates, which were 255 and 235.3, 255 and 235.5 respectively. Combined with the characteristics of these areas, it is concluded that these areas have a high flow of people or concentrated business districts. This phenomenon has also been verified in combination with the urban size effect theory of Bettencourt et al., where people with high traffic or rich commercial clusters attract thieves. This suggests that the area may have different conditions or characteristics than other regions, such as high traffic or rich business clusters that attract thieves and become a crime hotspot [@bettencourt2007growth].
